SiO Molecular Jets around young stars - A numerical perspective

### 2013-10-14 PAPER/Paola_1/ms_ver2.tex
Modified bib file and added Paola comments and ms_ver2.tex which incorporates the comments

\newcommand{\figpath}{PFIGS/}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Further, the velocities obtained from the dynamical simulations are in fact the
jet flow velocities. They can well be different from the shock
velocities depending upon the density contrast between the jet and
ambient medium. For the present study, we will consider two
cases. Firstly, the most simple one in which the jet flow velocity is
same as the shock velocity. Considering that there is a density
contrast in the simulation setup this assumption is will
overestimate the shock velocity. In order to correct the same we also
consider a case in which the shock speed of a dense plug accumulated
between the jet and medium is estimated by balancing the the respective ram
pressures \citep{Masson:1993p9661},
\begin{equation}
\label{eq:shockjetvel}
v_{\rm s} = \frac{v_{\rm j}}{1 + \eta^{-0.5}} = \frac{v_{\rm j}}{\delta}
\end{equation}
where v$_{\rm s}$ is the velocity of dense plug and v$_{\rm j}$ is the
jet velocity. A comparative view of using such velocity corrections and
functional forms of SiO abundances are discussed in
Sec~\ref{ssec:sioabunem}.

@@

\section{Parameter Survey}
\label{sec:parasurvey}
The results and analysis for the present work are divided in two parts,
viz., the dynamical numerical simulations and the post-processing with
radiative transfer code. For each part we have used certain parameters
and the effect of changing them is studied. Ideally all these parameters should come from observational
results. However, not all quantities needed for our study are well
constrained by observations, thus allowing us vary them as
free parameters. Such a parameter survey provides better handle on the range of allowed
values on qualitative comparison with observations. 
%

@@

A value of $\eta > 1$, means that
that jet is over dense with respect to the medium. In all our runs with
different cooling prescriptions, we have assumed the jet to be
over-dense by 10 times that of the ambient medium. Additionally, for
the atomic, tabulated and molecular cooling runs we have also used a
value of $\eta$ of the order of unity indicating similar densities in
the jet and the medium. 
The magnetic field strength is kept to be fixed using $\beta = 10.0$,
for all our runs. Table~\ref{tab:result1} lists all the runs with
varying $\eta$ and cooling prescriptions, along with the peak
intensity and line widths at the bow shock.
%

To obtain the SiO emission maps and corresponding
spectra, two additional free parameters are required along with other
inputs described in section~\ref{sec:radtrans} . They are
the fractional abundance profile of SiO and the angle of inclination with
respect to line of sight denoted by $\phi$. Section~\ref{ssec:sioabun}
describes all the profiles used for the present study. The obvious
choice for the inclination angle, $\phi$ is 90$^{\circ}$ indicating
the outflow is in the plane of the sky. 
Additionally, we have used two other angles of inclination apart from the
plane of the sky, i.e, $\phi = 45^{\circ}, 60^{\circ}$ to compare
results with observations. The runs with different fractional
abundance profiles are described in table~\ref{tab:result2}. A
parameter introduced here is $\delta$ which is essentially the ratio
of shock velocity v$_{\rm s}$ and physical jet velocity v$_{\rm
  j}$ such that its value depends on density contrast $\eta$
(see Eq.~\ref{eq:shockjetvel}).

@@

The reference run in our calculation has density contrast $\eta$ = 3
with a plasma beta, $\beta$ = 10. The jet for this run is 89\% atomic,
10\% molecular and 1\% ionized to begin with. This jet enters the
ambient medium with a velocity of v$_{\rm jet,0}$ = 100 km
s$^{-1}$. The cooling in the jet is
via molecular cooling prescription, whereby the hydrogen fractions are
evolved with cooling contributions for abundant molecules like CO, OH
etc. The final state (at time $\sim$ 1000 yr) density, temperature and velocity obtained from this 
run is post-processed to obtain the observational features
corresponding to SiO molecule. 
For the radiative transfer calculation, the source in the reference
run is stationary and the jet is in plane of the sky ($\phi =
90^{\circ}$). A gaussian profile is used for the fractional abundance
of SiO.
%TheSiO fractional abundance used is that obtained
%from equation~\ref{eq:fracSio}. 
Further, the shock velocity which determines
the fractional abundance, v$_{\rm s}$, is less than the jet velocity in the
flow, such that their ratio $\delta$ is given by 1./$(1.0 +
\eta(z)^{-0.5})$, where $\eta(z)$ is the density contrast as a
function of height from the base of the jet, and is inversely proportional to the ambient density at the same height.

@@

\section{ALMA view}
\label{sec:ALMAview}
%
In order to see the interior structure of the molecular outflow, high
resolution interferometric observations are needed. To demonstrate how
jets similar to those we have modeled would look through such
observations we have performed synthetic Atacama Large Millimeter
Array (ALMA) observations using the Common Astronomy Software
Applications package (CASA). The output from the radiative transfer
were scaled to 300pc and used as the sky models for observations of
the J=2-1, J=5-4 and J=8-7 lines at frequencies of 86.85, 217.10 and
347.33$\,$GHz, which fall in ALMA bands 3, 6 and 7. The observations
were simulated using cycle 1 ALMA for 30 minutes total integration
time, in the configurations c32-6, c32-4 and c32-3 for the three
lines, giving beam sizes of 0.72'', 0.47'' and 0.58'' and
sensitivities of 0.05, 0.07 and 0.09$\,$mJy/beam for the three lines
(as calculated by the ALMA online sensitivity calculator) with
velocity resolutions of $\sim$1$\,$km/s.
 Figure~\ref{fig:almafig}
presents the AMLA prediction obtained from our reference run. The
left panel shows the integrated emission of SiO J = 2-1 line while the right
shows the PV diagram for the same line and over laid are
the contours of SiO J = 5-4 and J = 8-7. The contour levels are marked
on the color bar of each panel.

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/ALMAmaps3.png}
 \caption{{\bf Left:} The integrated intensity map of the three SiO lines simulated. The heat map shows the 5-4 line intensity (in units of Jy$\,$kms$^{-1}$/beam), the blue contours show the J=8-7 line intensity and the green contours show the J=2-1 line intensity. {\bf Right:} The PV diagram taken along the axis of the jet (in units of Jy/beam), showing the higher J transitions highlighting the knots of the jet and broad emission at the bow shock. In both panels the ticks on the color bar represent the contour levels, the levels not shown on the integrated intensity color bar are double the previous level.}
\label{fig:almafig}
\end{figure*}

ALMA Cycle 0 observations have recently reported complex kinematic
features of young bipolar molecular outflows. In particular,
\cite{Arce:2013p14902} observed morphology, kinematics and entrainment
mechanism of the HH 46/47 molecular outflow from a low mass nearby
source (d = 450 pc) using CO(1-0). They find that the red
lobe of the outflow exhibits a very complex structure with a
collimated episodic wind whereas the outflow in the blue lobe is consistent with
wide-angle wind model. Three major clumps in the red lobe along the
outflow axis show a distinctive signature in emission maps. In
addition, these clumps appear as {\em{wedges}} in the PV
diagram. The observed {\em{wedges}} are very similar in structure to that 
seen in the right panel of figure~\ref{fig:almafig}. A quantitative
comparison is not possible due to obvious difference of molecular
lines. Another source which is studied in details with ALMA band 7 is
one of the most energetic, luminous molecular outflow
G331.512-0.103 \cite{Merello:2013p15066}. Observations of this outflow
associated with a high mass star reveal a compact, extremely young
bipolar outflow and a symmetric outflowing shell. Further, their PV
diagrams with SiO and H$^{13}$CO$^{+}$ show a characteristic peak at
-45\,km\,s$^{-1}$ along the primary axis that could associated with a 
molecular {\em bullet} of high-velocity dense material as predicted from our
synthetic maps.

\section{Discussion}
\label{sec:discussion}
Here we compare the various results with SiO observation of young low
mass and high mass outflows.

\subsection{EHV component and Molecular bullets}
\label{ssec:EHV}
Multiline mm-wavelength surveys of SiO emission towards a sample of molecular
outflows have shown that these outflows exhibit both low and high
velocity SiO emission possibly arising from two distinct regimes. The slow
component in the case of a typical outflow, L1448, is believed to arise from
the shell of accelerated ambient gas with SiO abundance
lower by two orders of magnitude than regions emitting the high velocity component.
\cite{Codella:1999p12584}. In addition, an interesting feature of EHV
gas is seen in many young molecular outflows. The origin of EHV
component is still a mystery, however, recent high resolution
observations have tried to study the composition of such extremely high
velocity gas \cite{Tafalla:2010p14759}. Typically, EHV gas feature in
high density gas tracers (like CO and SiO) 
and show significantly weak emission of $\sim$ 0.1\,K.
%

In IRAS 04166+2706, \cite{SantiagoGarcia:2009p13972} have shown 
that the EHV CO(2-1) emission is mapping a jet-like
feature that runs along the middle and consists of a collection of
discrete peaks. These peaks of EHV emission are placed located in a
perfect symmetry on both sides of the outflow lobe. This symmetry
indicate that such EHV peaks might arise from events that took place
near the central source and have since propagated in the flow
\cite{Bachiller:1990p11196, Tafalla:2011p14051}. The dynamical 
model in conjunction with proper radiative transfer calculations presented here 
agrees very well with the above scenario. The {\em{bullets}} in our
work are injected into the flow in a sinusoidal manner along with a
collimated atomic jet. These pulsating ejections do interact with the medium
via shocks and exciting high velocity gas. Cooling associated with
molecules and tracing of H$_{2}$ chemistry with the flow allows us to
consistently identify the regions where molecular hydrogen is formed
and disassociated (see fig.~\ref{fig:Hmolform}). The SiO emission that we
obtain from our radiative transfer calculations is associated with
regions where shocked H$_{2}$ gas is present. A contour map of SiO
emission obtain from the reference run is shown in
figure~\ref{fig:ltconts}. Such a model can very well explain the
shocked SiO velocity components observed in L1157 molecular outflow \cite{Gueth:1998p14058}. 
%

Further, we find SiO emission coming from velocities 
of the order of 40-60 km\,s$^{-1}$ (see figs.~\ref{empvspec60} and
~\ref{empvspec45}). Such a velocity range is typically associated with
EHV gas seen in majority of the outflows \cite{Tafalla:2011p14051}. Additionally,
the EHV peaks show a distinct sawtooth pattern in the PV diagram
\cite{SantiagoGarcia:2009p13972}. Such a sawtooth pattern is very
similar to the pattern that we see in PV diagrams from
our models (see figs.~\ref{empvspec60} and ~\ref{empvspec45}). 
All of the above striking similarities from observations of EHV gas
and synthetic spectra and PV diagrams gives a very formidable backing
to the idea that such an emission could arise due to shock interactions of
internal knots in the flow. In the section below we further compare
the line ratios and intensities from our results with single dish
measurements of typical outflows with EHV emission.

\subsection{Line transitions and ratio}
Numerous molecular outflows with jet-like bullets have been
observationally studied till date. In particular, H7-11
\citep{Bachiller:1998p14725}, IRAS 04166+2706
\citep{SantiagoGarcia:2009p13972, Tafalla:2010p14759}, HH211 \citep{Nisini:2002p14418}
L1448 \citep{Bachiller:1991p14732,Nisini:2007p13128,
  Tafalla:2010p14759} and L1157 \citep{Nisini:2007p13128} have
shown clear signatures of EHV component. Among them L1448, HH211 and
L1157 have been studied in details using an extensive multi-line
survey of SiO. Such a multi-frequency analysis helps to understand the dependence of
excitation conditions for these lines on velocity, based on existing shock
models.  
%

Figure~\ref{fig:ltconts} show symmetrical contour maps of integrated 
emission convolved with a 2\arcsec beam coming from different line
transitions for our reference model. The most
striking feature seen in that figure is the progressive shift of
emitting region from the interface to internal knots with increase in
excitation energies of different lines. In particular, the lowest
transition J=1-$>$0 shows most of the emission coming from the
interface between the jet and ambient medium, along with emission from
the dense knots formed at the base of the flow. While, emission coming
due to high energy transitions, J=5-$>$4 and 8-$>$7, 
is more concentrated in the inner jet regions and arise mainly from
the shocks due to internal knots. Such a trend in emission with excitation energies coming from different SiO line transitions
have been observed in many young outflows (for e.g., {\it{L1448}}
bullets \citep{Nisini:2007p13128}, {\it{HH211}}
\citep{Chandler:2001p14376, Nisini:2002p14418,
  Hirano:2006p14411}). Further, the evolved post-shock gas near the
primary bow shock also show bright emission for these high J transitions. 
This gas is linked to the thermal instability associated with cooling flows (see
section~\ref{ssec:coolres}). The sub-structure seen close to the
primary bow shock with high resolution does give a sense of clumpiness
in the flow backing the suggestions to explain clumpy SiO emission by
\citealt{Chandler:2001p14376}.
%

@@

\cite{Nisini:2007p13128} have shown that the current
plane-parallel shock models fit reasonably well the physical
conditions traced by the SiO emission i.e., temperature $\lesssim$
1000\,K and H$_{2}$ number density $\sim$ 10$^{5}$-10$^{6}$
cm$^{-3}$. They can also fit the observed fractional SiO abundance of
$\sim$ 10$^{-7}$ in these outflows. However, they fail to predict all
the line profiles and in particular their similarities for low and
high excitation lines. Line emission for SiO from our model also traces similar
physical environment. Fig.~\ref{fig:Hmolform} shows that regions close
to the internal knot have temperatures up to 1000\,K and molecular
hydrogen density of the order of 10$^{6}$ cm$^{-3}$ (also see fig.~\ref{grid}).
These regions are where bulk of the
SiO emission is obtained in our models. Additionally, the internal
knots move with velocity of 70-90 km\,s$^{-1}$, indicating a
fractional SiO abundance between 10$^{-6}$ to 10$^{-8}$, depending on the
choice of abundance profile (see fig~\ref{abun}). 
The line profile shapes obtained from this study fit reasonable well with that observed by
\cite{Nisini:2007p13128}, especially for the bullets seen in
L1448. Interestingly, we see in our models, where we relax the plane parallel
approximation, the line profiles are equally broadened for low as well
as high excitation lines. The top panel of fig.~\ref{fig:lineratio} shows spectra for three line
transitions for the reference run but with an angle of inclination,
$\phi$ = 60$^{\circ}$ and convolved with a single dish beam of
15\arcsec. Additionally, the predicted integrated
intensities also lie within a factor of two from the values of L1448 bullets
observed with JCMT and IRAM (Table 2 of \citealt{Nisini:2007p13128}  and
table~\ref{tab:result2} of this work).
%

The variation of integrated single dish emission with upper transition
levels J$_{\rm up}$ for different abundance profiles is shown in the
right panel of fig.~\ref{fig:lineratio}. It is clear, that a emission
obtained using a top hat profile is brighter as compared to one
obtained with a gaussian profile, due to its obvious high estimate of
SiO abundance in regions of interest. However, curves obtained from
both these profiles show a very similar shape of a distinct rise followed by a fall in
integrated emission for higher transitions with a peak in
emission for J$_{\rm up}$ = 3. Similar profiles are also predicted and
used to estimate the physical conditions assuming optically
thin emission from LVG modeling for L1448 and L1157
\cite{Nisini:2007p13128}. However, this usual approach did not give
accurate results for line ratios SiO(8-7)/(5-4) and SiO(5-4)/(2-1)
in case of HH212. Values $\approx$ 1 for both ratios were
achieved for low mass in LTE regime \citep{Cabrit:2007p13804, Lee:2008p13697}. 
and also for molecular outflows from massive
young stellar object IRAS 17233-3606 \cite{Leurini:2013p13165}.
These line ratios from the present
non-LTE radiative transfer model are shown in the bottom left panel of
fig.~\ref{fig:lineratio}. Their values lie very close to unity as
indicated from observations. 
%Further, we have estimated the optical
%depth in regions emitting SiO and find them to be optically thick with
%an optical depth $\tau \sim 10$. Thus, assuming optically thin SiO emission
%from molecular outflows especially from shocked dense knots would
%only provide upper limits. 
In summary, the shape, ratio and peak intensity of the spectra
obtained from our model fit very well to observed values implying
that SiO emission from our model is tracing the regions with right physical
conditions that are required to emit SiO in gas phase via shocks.

@@

\subsection{Slow Component and Grain Chemistry}
%
We have presented a formidable model to explain the EHV component seen
in SiO emission from several young outflows. The reference model
consistently solves for the H$_{2}$ chemistry in presence of
appropriate cooling. The steady state temperature, density and velocities obtained from dynamical
simulation is post-processed to obtain SiO maps, spectra and PV
diagrams. One essential ingredient required for the radiative transfer modeling
is the SiO fractional abundance and its dependence on shock velocity. 
Though 1D models that study the formation of SiO in gas phase from grain-grain and grain-mantle
collisions exists, their focus is mainly on weaker shocks
\citep{Schilke:1997p14140, Caselli:1997p14853, Gusdorf:2008p13800}. 
The pulsating jet propagation model presented here have shock speeds
reaching up to 100 km\,s$^{-1}$. For simplicity we have used two basic
prescription of SiO abundance profile as a function of shock
velocity. Though the shape of the profile is uncertain the upper and
lower limits used are backed by observational evidences. Ideally for a consistent
dependence of SiO abundance on shock velocity, one would have to also
solve for silicon and SiO chemistry in a manner similar to that of
H$_{2}$. Further by evolving the dynamical simulations with complex
grain chemistry for more than 10$^{4}$ years (time to destroy SiO in
gas phase \cite{Codella:1999p12584}),  one can trace the history of 
SiO in molecular outflows. However, such a numerically expensive model 
with complex chemistry is beyond the scope of this paper.  
%

>>>>

\newcommand{\figpath}{/Users/bhargavvaidya/MyProject/work/Leeds_Uni/SiOJets_New/PAPER/PFIGS/}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Further, the velocities obtained from the dynamical simulations are in fact the
jet flow velocities. They can well be different from the shock
velocities depending upon the density contrast between the jet and
ambient medium. For the present study, we will consider two
cases. Firstly, the most simple one in which the jet flow velocity is
same as the shock velocity. Considering that there is a density
contrast in the simulation setup this assumption is will
overestimate the shock velocity. In order to correct the same we also
consider a case in which the shock speed of a dense plug accumulated
between the jet and medium is estimated by balancing the the respective ram
pressures \citep{Masson:1993p9661},
\begin{equation}
\label{eq:shockjetvel}
v_{\rm s} = \frac{v_{\rm j}}{1 + \eta^{-0.5}} = v_{\rm j}\delta
\end{equation}
where v$_{\rm s}$ is the velocity of dense plug and v$_{\rm j}$ is the
jet velocity. A comparative view of using such velocity corrections and
functional forms of SiO abundances are discussed in
Sec~\ref{ssec:sioabunem}.

@@

\section{Parameter Survey}
\label{sec:parasurvey}
The results and analysis for the present work are divided in two parts,
viz., the dynamical numerical simulations and the
radiative transfer code. For each part we have used certain parameters
and the effect of changing them is studied. Ideally all these parameters should come from observational
results. However, not all quantities needed for our study are well
constrained by observations, thus they are considered as
free parameters. Such a parameter survey provides better handle on the range of allowed
values on qualitative comparison with observations. 
%

@@

A value of $\eta > 1$, means that
that jet is over dense with respect to the medium. In all our runs with
different cooling prescriptions, we have assumed the jet to be
over-dense by 10 times compared to the ambient medium. Additionally, for
the atomic, tabulated and molecular cooling runs (see section~\ref{sec:chem}
for description of different cooling prescriptions) we have also used a
value of $\eta$ of the order of unity indicating similar densities in
the jet and the medium. 
The magnetic field strength is kept to be fixed using $\beta = 10.0$,
for all our runs. Table~\ref{tab:result1} lists all the runs with
varying $\eta$ and cooling prescriptions, along with the peak
intensity and line widths at the bow shock.
%

To obtain the SiO emission maps and corresponding
spectra, two additional free parameters are required along with other
inputs described in section~\ref{sec:radtrans} . They are
the fractional abundance profile of SiO and the angle of inclination with
respect to line of sight denoted by $\phi$. Section~\ref{ssec:sioabun}
describes all the profiles used for the present study. $\phi$ = 90$^{\circ}$ indicates that
the outflow is in the plane of the sky. 
Additionally, we have used two other angles of inclination apart from the
plane of the sky, i.e, $\phi = 45^{\circ}, 60^{\circ}$ to compare
results with observations. The runs with different fractional
abundance profiles are described in table~\ref{tab:result2}. A
parameter introduced here is $\delta$ which is essentially the ratio
of shock velocity v$_{\rm s}$ and physical jet velocity v$_{\rm
  j}$ such that its value depends on density contrast $\eta$
(see Eq.~\ref{eq:shockjetvel}).

@@

The reference run in our calculation has density contrast $\eta$ = 3
with a plasma beta $\beta$ = 10. The jet for this run is 89\% atomic,
10\% molecular and 1\% ionized to begin with. This jet enters the
ambient medium with a velocity of v$_{\rm jet,0}$ = 100 km
s$^{-1}$. The cooling in the jet is
via molecular cooling prescription, whereby the hydrogen fractions are
evolved with cooling contributions for abundant molecules like CO, OH
etc. The final state (at time $\sim$ 1000 yr), i.e, density, temperature and velocity obtained from this 
run are given in input to the radiative transfer code to obtain the observational features
corresponding to SiO molecule. 
For the radiative transfer calculation, the source in the reference
run is stationary and the jet is in the plane of the sky ($\phi =
90^{\circ}$). A gaussian profile is used for the fractional abundance
of SiO.
%TheSiO fractional abundance used is that obtained
%from equation~\ref{eq:fracSio}. 
Further, the shock velocity which determines
the fractional abundance, v$_{\rm s}$, is less than the jet velocity in the
flow, such that their ratio $\delta$ is given by
equation~\ref{eq:shockjetvel}, where $\eta$ is the density contrast as a
function of height from the base of the jet.

@@

\section{ALMA view}
\label{sec:ALMAview}
%
In order to see the interior structure of the molecular outflow and
test the model predictions, high
resolution interferometric observations are needed. 
We have performed synthetic Atacama Large Millimeter
Array (ALMA) observations using the Common Astronomy Software
Applications package (CASA). The output from the radiative transfer
were scaled to a distance of 300pc and used as the sky models for observations of
the J=2-1, J=5-4 and J=8-7 lines at frequencies of 86.85, 217.10 and
347.33$\,$GHz, which fall in ALMA bands 3, 6 and 7, respectively. The observations
were simulated using cycle 1 ALMA for 30 minutes total integration
time, in the configurations c32-6, c32-4 and c32-3 for the three
lines, giving beam sizes of 0.72'', 0.47'' and 0.58'' and
sensitivities of 0.05, 0.07 and 0.09$\,$mJy/beam for the three lines
(as calculated by the ALMA online sensitivity calculator) with
velocity resolutions of $\sim$1$\,$km/s.
 Figure~\ref{fig:almafig}
presents the ALMA prediction obtained from our reference run. The
left panel shows the integrated emission of SiO J = 2-1 line while the right
shows the PV diagram for the same line and over laid are
the contours of SiO J = 5-4 and J = 8-7. The contour levels are marked
on the color bar of each panel.

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/ALMAmaps3.png}
 \caption{{\bf Left:} The integrated intensity map of SiO(2-1), (5-4)
   and (8-7). The heat map shows the 2-1 line intensity (in units of Jy$\,$kms$^{-1}$/beam), the blue contours show the J=8-7 line intensity and the green contours show the J=5-4 line intensity. {\bf Right:} The PV diagram taken along the axis of the jet (in units of Jy/beam), showing the higher J transitions highlighting the knots of the jet and broad emission at the bow shock. In both panels the ticks on the color bar represent the contour levels.}
\label{fig:almafig}
\end{figure*}

ALMA Cycle 0 observations have recently reported complex kinematic
features of young bipolar molecular outflows. In particular,
\cite{Arce:2013p14902} observed morphology, kinematics and entrainment 
of the HH 46/47 molecular outflow from a low mass nearby
source (d = 450 pc) using CO(1-0). They find that the red
lobe of the outflow exhibits a very complex structure with a
collimated episodic wind, whereas the outflow in the blue lobe is
consistent with a
wide-angle wind model. Three major clumps in the red lobe along the
outflow axis appear as {\em{wedges}} in the PV
diagram (similar to that seen in the right panel of fig~\ref{fig:almafig}). 
A quantitative comparison is not possible due to the different
molecule observed. Another source which is studied in details with ALMA band 7 is
one of the most energetic, luminous molecular outflows
G331.512-0.103 (\citealt{Merello:2013p15066}). Observations of this outflow
associated with a high mass star reveal a compact, extremely young
bipolar outflow and a symmetric outflowing shell. Further, their PV
diagrams with SiO and H$^{13}$CO$^{+}$ show a characteristic peak at
-45\,km\,s$^{-1}$ along the primary axis that could associated with a 
molecular {\em bullet} of high-velocity dense material as also found in our
synthetic maps at inclination angles of 60$^{\circ}$ and 45$^{\circ}$.

\section{Discussion}
\label{sec:discussion}
Here we compare our results with SiO observations of young low
mass and high mass outflows.

\subsection{EHV component and Molecular bullets}
\label{ssec:EHV}
Multiline mm-wavelength surveys of SiO emission towards a sample of molecular
outflows have shown that these outflows exhibit both low and high
velocity SiO emission possibly arising from two distinct regimes. The slow
component in the case of a typical outflow, L1448, is believed to arise from
the shell of accelerated ambient gas with SiO abundance
lower by two orders of magnitude than regions emitting the high
velocity component
\cite{Codella:1999p12584}. In addition, an interesting feature of EHV
gas is seen in many young molecular outflows. The origin of the EHV
component is still a mystery, however, recent high resolution
observations show that EHV gas is seen in
high density gas tracers (like CO and SiO) and the line have
intensities $\sim$ 0.1 K (\citealt{Tafalla:2010p14759}).
%

In IRAS 04166+2706, \cite{SantiagoGarcia:2009p13972} have shown 
that the EHV CO(2-1) emission is mapping a jet-like
feature that consists of a collection of
discrete peaks symmetrically placed on both outflow lobes. This symmetry
indicate that such EHV peaks might arise from events that took place
near the central source and have since propagated in the flow
\citep{Bachiller:1990p11196, Tafalla:2011p14051}. The dynamical 
model in conjunction with proper radiative transfer calculations presented here 
agrees very well with the above scenario. The {\em{bullets}} in our
work are injected into the flow in a sinusoidal manner along with a
collimated atomic jet. These pulsating ejections do interact with the medium
via shocks and exciting high velocity gas. Including the cooling associated with
molecules and the H$_{2}$ chemistry allows us to
consistently identify the regions where molecular hydrogen is formed
and disassociated (see fig.~\ref{fig:Hmolform}). The SiO emission that we
obtain from our radiative transfer calculations is associated with
regions where shocked H$_{2}$ gas is present. A contour map of SiO
emission obtained from the reference run is shown in
figure~\ref{fig:ltconts}. Such a model can very well explain the
shocked SiO velocity components observed in the L1157 molecular outflow \cite{Gueth:1998p14058}. 
%

Further, we find SiO emission coming from velocities 
of the order of 40-60 km\,s$^{-1}$ (see figs.~\ref{empvspec60} and
~\ref{empvspec45}). Such a velocity range is typically associated with
EHV gas seen in the majority of the outflows (\citealt{Tafalla:2011p14051}). Additionally,
the EHV peaks show a distinct sawtooth pattern in the PV diagram
(\citealt{SantiagoGarcia:2009p13972}). Such a sawtooth pattern is very
similar to the pattern that we see in PV diagrams from
our models (see figs.~\ref{empvspec60} and ~\ref{empvspec45}). 
All of the above striking similarities from observations of EHV gas
and synthetic spectra and PV diagrams gives a very formidable backing
to the idea that such an emission could arise due to shock interactions of
internal knots in the flow. In the section below we further compare
the line ratios and intensities from our results with single-dish
measurements of typical outflows with EHV emission.

\subsection{Line transitions and ratio}
Numerous molecular outflows with jet-like bullets have been
observationally studied till date. In particular, H7-11
\citep{Bachiller:1998p14725}, IRAS 04166+2706
\citep{SantiagoGarcia:2009p13972, Tafalla:2010p14759}, HH211 \citep{Nisini:2002p14418}
L1448 \citep{Bachiller:1991p14732,Nisini:2007p13128,
  Tafalla:2010p14759} and L1157 \citep{Nisini:2007p13128} have
shown clear signatures of the EHV component. Among them, L1448, HH211 and
L1157 have been studied in detail using an extensive multi-line
survey of SiO. Such a multi-frequency analysis helps to understand the dependence of
excitation conditions for these lines on velocity, based on existing shock
models.  
%

The most striking feature seen in the figure~\ref{fig:ltconts} is the progressive shift of
emitting region from the interface to internal knots with increase in
excitation energies of different lines. In particular, the lowest
transition J=1-$>$0 shows most of the emission coming from the
interface between the jet and ambient medium, along with emission from
the dense knots formed at the base of the flow. On the other hand, SiO (5-4) and (8-7) emission, 
is more concentrated in the inner jet regions and arise mainly from
the shocks due to internal knots. Such a trend in emission with excitation energies coming from different SiO line transitions
have been observed in young outflows (for e.g., {\it{L1448}}
bullets \citep{Nisini:2007p13128}, {\it{HH211}}
\citep{Chandler:2001p14376, Nisini:2002p14418,
  Hirano:2006p14411}. Further, the evolved post-shock gas near the
primary bow shock also show bright emission for these high J transitions. 
This gas is linked to the thermal instability associated with
radiative jets (see section~\ref{ssec:coolres}). The sub-structure seen close to the
primary bow shock with high resolution does give a sense of clumpiness
in the flow backing the suggestions to explain clumpy SiO emission has
also been observed by \cite{Chandler:2001p14376}.
%

@@

\cite{Nisini:2007p13128} have shown that the current
plane-parallel shock models fit reasonably well the physical
conditions traced by the SiO emission i.e., temperature $\lesssim$
1000\,K and H$_{2}$ number density $\sim$ 10$^{5}$-10$^{6}$
cm$^{-3}$. They can also fit the observed fractional SiO abundance of
$\sim$ 10$^{-7}$ in these outflows. However, they fail to predict all
the line profiles and in particular their similarities for low and
high excitation lines. Line emission for SiO from our model also traces similar
physical environment. Fig.~\ref{fig:Hmolform} shows that regions close
to the internal knot have temperatures up to 1000\,K and molecular
hydrogen density of the order of 10$^{6}$ cm$^{-3}$ (also see
fig.~\ref{grid}). These are the regions where the bulk of the SiO
emission is present, with fractional abundances between 10$^{-6}$ to 10$^{-8}$.
The line profiles shapes obtained from our study fit reasonable well with that observed by
\cite{Nisini:2007p13128}, especially for the bullets seen in
L1448. Interestingly, in our models, the line profiles are equally broadened for low as well
as high excitation lines. The top panel of fig.~\ref{fig:lineratio} shows spectra for three line
transitions for the reference run but with an angle of inclination
$\phi$ = 60$^{\circ}$ and convolved with a single dish beam of
15\arcsec. Additionally, the predicted integrated
intensities also lie within a factor of two from the values of L1448 bullets
observed with JCMT and IRAM (Table 2 of \citealt{Nisini:2007p13128}  and
table~\ref{tab:result2} of this work).
%

The variation of integrated single dish emission with upper transition
levels J$_{\rm up}$ for different abundance profiles is shown in the
right panel of fig.~\ref{fig:lineratio}. It is clear that the emission
obtained using a top hat profile is brighter than the one obtained
with a Gaussian SiO profile. However, curves obtained from
both these profiles show a very similar shape of a distinct rise followed by a fall in
integrated emission for higher transitions with a peak in
emission for J$_{\rm up}$ = 3. \cite{Nisini:2007p13128} found that
their homogeneous slab model coupled with LVG radiative transfer, did
not reproduce the observed line ratios SiO(8-7)/(5-4) and
SiO(5-4)/(2-1) of HH212. Values $\approx$ 1 for both ratios have been
found in for low mass assuming LTE regime \citep{Cabrit:2007p13804,
  Lee:2008p13697} and also for molecular outflows from massive
young stellar object IRAS 17233-3606 (\citealt{Leurini:2013p13165}).
These line ratios from the present
non-LTE radiative transfer model are shown in the bottom left panel of
fig.~\ref{fig:lineratio}. Their values lie very close to unity as
indicated from observations. 
%Further, we have estimated the optical
%depth in regions emitting SiO and find them to be optically thick with
%an optical depth $\tau \sim 10$. Thus, assuming optically thin SiO emission
%from molecular outflows especially from shocked dense knots would
%only provide upper limits. 
In summary, the shape, ratio and peak intensity of the spectra
obtained from our model fit very well to observed values implying
that SiO emission from our model is tracing the regions with right physical
conditions that are required to emit SiO in the gas phase via shocks.

@@

\subsection{Slow Component and Grain Chemistry}
%
We have presented a model able to explain the spectral characteristics
of the SiO EHV component seen
in several young outflows. The reference model
consistently solves for the H$_{2}$ chemistry in presence of
appropriate cooling. The steady state temperature, density and velocities obtained from dynamical
simulation is used as input in the non-LTE radiative transfer code to obtain SiO maps, spectra and PV
diagrams. One essential ingredient required for the radiative transfer modeling
is the SiO fractional abundance and its dependence on shock velocity. 
Though 1D models that study the formation of SiO in gas phase from grain-grain and grain-mantle
collisions exists, their focus is mainly on weaker shocks
\citep{Schilke:1997p14140, Caselli:1997p14853, Gusdorf:2008p13800}. 
The pulsating jet propagation model presented here have shock speeds
reaching up to 100 km\,s$^{-1}$. We have used two
simplified SiO abundance profiles as a function of shock
velocity. Though the shape of the profile is uncertain, the upper and
lower limits used are backed by observational evidences. Ideally, for a consistent
dependence of SiO abundance on shock velocity, one would have to also
solve for silicon and SiO chemistry in a manner similar to that of
H$_{2}$. However, such a numerically expensive model 
with complex chemistry is beyond the scope of this paper 
as this could imply to include dust grains and sputtering 
as well as grain-grain collisions (i.e. an MHD multi-fluid approach).  
%

The model presented in this paper targets the very early phases of molecular
outflows i.e., dynamical time scales of 1000 years.
As discussed in sections above, the unique combination of pulsating
jet with chemistry and cooling followed by non-LTE radiative transfer
calculations has shown success in fitting the line shapes, integrated intensity and line ratios of SiO
transitions arising from EHV gas. However, the synthetic spectra show
no signatures of slow SiO component that accompanies EHV emission seen
in many outflows \citep[e.g.,][]{Lefloch:1998p13983,
  Codella:1999p12584}. 
The physical mechanism for the origin of the slow SiO component is still a matter
of debate. \cite{Codella:1999p12584} suggest that it arises due to the slowing down of
shocked gas as they age. The time scale estimated for the shocked gas
to slow down is of the same order of magnitude as the SiO destruction
time scale i.e., 10$^{4}$ years (\citealt{Codella:1999p12584}). In this
case, we do not expect to see the slow velocity SiO component, as our
models stop at $\sim$ 1000 yr. However, we do see initial 
hints of slowing down of more evolved gas in terms of the velocity
shift of about 7-10 km\,s$^{-1}$ in peak emission between that
coming from the freshly formed internal knots close to the base of the
flow and that from the more evolved gas near the
primary bow shock. To ascertain this fact in more details, one would
need to track the primary bow shock for 10$^{4}$ years
using a larger simulation box and possibly with adaptive gridding 
which will be considered for future simulations. Such a long term
evolution will also help to provide a numerical model for the formation of HH objects
which are believed to be slowed down and more evolved forms of our
young molecular bullets, as suggested by \cite{Norman:1979p14858} and
  \cite{Hartigan:1987p11178}.
%
